---
title: "BayesGP: Fitting sGP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BayesGP: Fitting sGP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", fig.height=3, fig.width=5, margins=TRUE
)
knitr::knit_hooks$set(margins = function(before, options, envir) {
  if (!before) return()
  graphics::par(mar = c(1.5 + 0.9, 1.5 + 0.9, 0.2, 0.2), mgp = c(1.45, 0.45, 0), cex = 1.25, bty='n')
})
```


```{r setup}
options(warn=-1)
library(BayesGP)
```

# Fitting sGP

In this tutorial, we will use the `lynx` dataset as an example, which
can be accessed directly from R. Let’s load the dataset and visualize
it:

```{r}
data <- data.frame(year = seq(1821, 1934, by = 1), y = as.numeric(lynx))
data$x <- data$year - min(data$year)
plot(lynx)
```

Based on a visual examination of the dataset, we can observe an obvious
10-year quasi-periodic behavior in the lynx count with evolving
amplitudes over time. Therefore, we will consider fitting the following
sGP model:

$$
\begin{aligned}
  y_i|\lambda_i &\sim \text{Poisson}(\lambda_i) ,\\
  \log(\lambda_i) &= \eta_i = \beta_0 + g(x_i) + \xi_i,\\
  g &\sim \text{sGP} \bigg(a = \frac{2\pi}{10}, \sigma\bigg),\\
  \xi_i &\sim N(0,\sigma_\xi).
\end{aligned}
$$

Here, each $y_i$ represents the lynx count, $x_i$ represents the number
of years since 1821, and $\xi_i$ is an observation-level random
intercept to account for overdispersion effect.


## Prior Elicitation:

To specify the priors for the sGP boundary conditions and the intercept
parameter, we assume independent normal priors with mean 0 and variance
1000. For the overdispersion parameter $\sigma_\xi$, we assign an
exponential prior with $P(\sigma_\xi > 1) = 0.01$.

To determine the prior for the standard deviation parameter $\sigma$ of
the sGP, we employ the concept of predictive standard deviation (PSD).
We start with an exponential prior on the 50 years PSD:
$$P(\sigma(50)>1) = 0.01.$$ To convert this prior to the original
standard deviation parameter $\sigma$, we use the `compute_d_step_sGPsd`
function from the `sGPfit` package:

```{r}
prior_PSD <- list(u = 1, alpha = 0.01)
prior_SD <- BayesGP::prior_conversion_sgp(d = 50, prior = prior_PSD, a = 2*pi/10)
```

 
## Model Fitting:

To fit the sGP model with BayesGP, specify `model = "sGP"`, and then specify the frequency parameter `a` and the number of sB spline basis used to approximate `k`.

```{r}
mod <- model_fit(formula = y ~ f(x = year, 
                                    model = "sGP", 
                                    a = 2*pi/10, k = 30,
                                    sd.prior = list(prior = "exp", param = prior_SD, h = 2)) + 
                               f(x = x, 
                                    model = "IID", 
                                    sd.prior = list(prior = "exp", param = 0.5)),
                   
                 family = "Poisson",
                 data = data,
                 method = "aghq")
```

The fitted sGP model is presented below:

```{r}
summary(mod)
```

We can similarly use `predict` to evaluate the effect at different locations:

```{r}
pred_sGP <- predict(mod, variable = "year", newdata = data.frame(year = seq(min(data$year), max(data$year), by = 0.1)))
plot(mean ~ year, data = pred_sGP, type = 'l', col = 'black')
lines(q0.025 ~ year, data = pred_sGP, lty = 'dashed', col = 'grey')
lines(q0.975 ~ year, data = pred_sGP, lty = 'dashed', col = 'grey')
```

```{r}
mod <- model_fit(formula = y ~ f(x = year, 
                                    model = "sGP", 
                                    a = 2*pi/10, k = 30,
                                    sd.prior = list(prior = "exp", param = prior_SD, h = 2),
                                    boundary.prior = list(prec = c(Inf, Inf))) + 
                               f(x = x, 
                                    model = "IID", 
                                    sd.prior = list(prior = "exp", param = 0.5)),
                   
                 family = "Poisson",
                 data = data,
                 method = "aghq")
```


