---
title: "BayesGP: Fitting Model with Partial Likelihood"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BayesGP: Partial Likelihood}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", fig.height=3, fig.width=5, margins=TRUE
)
knitr::knit_hooks$set(margins = function(before, options, envir) {
  if (!before) return()
  graphics::par(mar = c(1.5 + 0.9, 1.5 + 0.9, 0.2, 0.2), mgp = c(1.45, 0.45, 0), cex = 1.25, bty='n')
})
```


```{r setup}
options(warn=-1)
library(BayesGP)
```

# Partial Likelihood Models in BayesGP

There are currently two implemented models in `BayesGP` that use the partial likelihood function for inference: the case-crossover model and the Cox Proportional Hazard (Coxph) model.



## Case-Crossover Model

With `BayesGP`, one can specify the argument `family` to `"cc"`, `"casecrossover"` or `"CaseCrossover"` to fit a case-crossover model.

Here we will use a simulated dataset:
```{r}
data <- as.data.frame(sim1data)
data$exposure <- data$exposure
mod <- model_fit(formula = case ~ f(x = exposure, 
                                    model = "IWP", 
                                    order = 2, k = 30,
                                    initial_location = median(data$exposure), 
                                    sd.prior = list(prior = "exp", param = list(u = 1, alpha = 0.5), h = 1)),
                 family = "cc",
                 strata = "subject",
                 weight = NULL,
                 data = data,
                 method = "aghq")

```

To take a look at its result:

```{r}
true_effect <- function(x) {3 *(x^2 - .5^2)}
plot(mod)
lines(I(true_effect(seq(0,1,by = 0.1)) - true_effect(median(data$exposure))) ~ seq(0,1,by = 0.1), col = "red")
```

Here the true effect used to simulate the data is shown as the red line. It is important to know that for case-crossover model, the intercept parameter and the `strata` level effects will not be identifiable.



## CoxPH Model


For Cox Proportional Hazard Model, one can specify the argument `family = "coxph"` to fit a CoxPH model with its partial likelihood.

Here we will illustrate with the `kidney` example from the `survival` package.

```{r}
data <- survival::kidney
head(data)
mod <- model_fit(formula = time ~ age + sex + f(x = id, 
                                    model = "IID", 
                                    sd.prior = list(prior = "exp", param = list(u = 1, alpha = 0.5))),
                 family = "coxph",
                 cens = "status",
                 data = data,
                 method = "aghq")
```

Take a look at the posterior for each fixed effect:

```{r}
samps_age <- sample_fixed_effect(mod, variables = "age")
samps_sex <- sample_fixed_effect(mod, variables = "sex")
par(mfrow = c(1,2))
hist(samps_age, main = "Samples for effect of age", xlab = "Effect")
hist(samps_sex, main = "Samples for effect of sex", xlab = "Effect")
```

 
